---
title: "Time series analysis"
author: "Villaseñor-Derbez, J.C."
date: "11 de febrero de 2016"
output: 
  html_notebook:
    code_folding: hide
    fig_caption: yes
    toc: yes
    toc_collapse: no
    toc_float: yes
---

# Part 1, Intro to TS

## First we load the packages that will be used:

```{r}
suppressPackageStartupMessages({
  library(forecast)
  library(tidyverse)
})
```

## Read the csv data and plot the Temperature agains month elasped

```{r}
arctic <- read.csv("ArcticTemp.csv")
plot(arctic$MonthElapse, arctic$Temp, type="l")
```

## convert to `ts` object

- Create a variable called `TempData` with just the temperature data
- Use `ts` to convert the data, specify the frequency of the cycle (12 for 12 months) and the year and month of first smapling
- `ts` objects can be directly plotted with base `plot`

```{r}
TempData <- arctic$Temp

TempTS <- ts(data=TempData, frequency=12, start=c(1975,1))

plot(TempTS)

```

## Decompose a time series

- A time series might have different components to it (seasonal, annual, decadal...), and we want to extract all that info
- Use `decompose` and base `plot` to look at the temporal trend, seasonal signature, and random noise in the data

```{r}
DecompTemp <- decompose(TempTS)

plot(DecompTemp)

```

## Forecasting

### Calculate parameters for existing data

- Use `HoltWinters` to "fit" the data
- Use base `plot` to see observed (black) and fitted (red) values

```{r}
Forecast <- HoltWinters(TempTS)
plot(Forecast)
```

### Forecast into the future

- Use what `HoltWinters` calculated to feed `forecast`
- `forecast` needs ou to specify how many points you want. In this case, we want 4 years so 4 Y * 12 M = 48 points.
- Use base `plot` to look at the projection

```{r}
Forecast2 <- forecast(Forecast, h=48)
plot(Forecast2)
```


# Part 2, Time series Analysis with ARIMA

In this case we have 2 time series. One for Residential use of energy and one for Industrial...

```{r, "ARIMA"}
#Load the data
EnergyUse <- read.csv("EnergyUse.csv")

#Convert data to time series format
ResidentialTS <- ts(EnergyUse$Residential, frequency=12, start=c(1993,1)) 
IndustrialTS <- ts(EnergyUse$Industrial, frequency=12, start=c(1993,1))

#Plot data
par(mfrow=c(2,1)) 
plot(ResidentialTS)
plot(IndustrialTS)

```

##Decompose data and plot it

### For Residential TS

```{r}
DecompRes <- decompose(ResidentialTS)
plot(DecompRes)
```

### For Industrial TS

```{r}
DecompInd <- decompose(IndustrialTS)
plot(DecompInd)
```

## We convert our data to stationary

- Just substract the temporal trend from the total data

```{r}
ResStationary <- DecompRes$x-DecompRes$trend
plot(ResStationary)
```

## Correlograms

- Time series data is often auto-correlated (yesterday's temp is a good predictor for today's temp...)
- We can inspect temporal correlations with a correlogram:

```{r}
ResACF <- acf(ResStationary[7:499])
```


#### Using auto.arima

```{r, "auto.arima"}
ResOpt=auto.arima(ResidentialTS) #This generates the parameters with a default of seasonal component
IndOpt=auto.arima(IndustrialTS, seasonal=FALSE) # This excludes the seasonal component
```

From the autoarima, we get that the values are:

Residential: ARIMA(1,0,1)(0,1,1)

Industrial: ARIMA(5,1,2)

And now we can use that to build an ARIMA

```{r, arima}

# We use arima to generate the model
ResARIMA=arima(ResidentialTS, order=c(1,0,1), seasonal=list(order=c(0,1,1)))
IndARIMA=arima(IndustrialTS, order=c(5,1,2))

# We can use histogram and qqplots to look for normality
hist(ResARIMA$residuals)
qqnorm(ResARIMA$residuals)

# And we use forecast to forecast our data
ForecastRes=forecast(ResARIMA, h=72)
ForecastInd=forecast(IndARIMA, h=72)

# And plot it
plot(ForecastRes)
plot(ForecastInd)

```