---
title: "Rstudio::conf(2018L)"
author: "Juan Carlos Villaseñor-Derbez"
date: "Jan 31 - Feb 4 2018"
output:
  html_document:
    code_folding: show
    fig_caption: yes
    toc: yes
    toc_collapse: no
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r}
knitr::opts_chunk$set(eval = FALSE)
```


## `tibbletime`

- `collapse_by` to copy or re-write time up. useful for then calculating `group_by` and `summarize`
- `rollify` is an adverb used to modify existing functions. `rollify(~mean(), window = 5)` calculates a 5-period running mean
- 

## [`infer`](https://github.com/andrewpbray/infer)

- Allen Downey

```{r}
data %>%
  specify() %>% #Define the model relation
  hypothesize() %>% # Specify null hypotheses
  generate() %>% # Create permutations of data
  calculate() %>%  # Calculate statistic
  visualize() # Get a plot of the distribution
```

## Tidy predictive modelling

- `broom`
- `tidyposterior`
  - Empirical validation
- `resample`
- `recipies`
- `yardstick`

## `tidyeval`

- Names of columns in the `aes` argument of a `ggplot` are "quoted", whereas the `data` argument is "evaluated"
- The `aes` uses non-standard evaluation
- `expr()` captures the expression that you give

```{r}
rlang::expr(a + b)
```

```{r}
(x1 <- rlang::expr(x + y))

(x2 <- rlang::expr(!!x1 + z))

(x3 <- rlang::expr(1 / !!x1))
```

```{r}
my_fun <- function(x, var){
  var <- enquo(var)
  
  y <- group_by(x, cyl) %>% 
    summarize(mean = mean(!!var, na.rm = T))
  
  return(y)
}
```

```{r}
my_fun(mtcars, disp)
```


## New operator in `zeallot`

`%<-%` Can be used to create n variables simultaneously. The syntax is `c(name1, name2, ..., namen) %<-% object`

The number of columns or vectors or numbers in `object` need to match `n`

```{r}
library(zeallot)
c(var1, var2) %<-% mtcars[, c(1, 4)]

var1
var2
```



 ## TensorFlow talk

- Packages:
  - `greta`:
    - simple and scalable statistical modelling in R
    - https://greta-dev.github.io/greta/
  - `tfruns`
  - [`keras`](https://t.co/BneOQfMe1V)
  - `tensorflow`
  - `cloudml`
  - `tfdeploy`
- Deep Learning with R (book)
- Deep Learning (book)
- https://tensorflow.rstudio.com/learn/gallery.html

## Google cloud stuff

- `dbplyr`
- `kerasjs`: export a keras model as javascript and embed it on an html document
- 

## Lesser know stars of the `tidyverse`

- `tibble::as_tibble`
- `tibble::tribble`
- `dplyr::na_if`
- `dplyr::select_if`
- `skimr::skim`
- `stringr::str_split`
- `tidyr::unnest`
- `forcats::fct_reorder`
- `forcats::fct_relevel`
- `reprex::reprex`
































