---
title: "Spatial Analysis with SF"
author: "Juan Carlos Villaseñor-Derbez"
date: "4 de octubre de 2017"
output: 
  html_notebook:
    code_folding: hide
    fig_caption: yes
    toc: yes
    toc_collapse: no
    toc_float: yes
---

# `SF` Package

## Load basic packages

```{r}
suppressPackageStartupMessages({
  library(sf)
  library(rgdal)
  library(raster)
  library(dplyr)
  library(fasterize)
  library(ggplot2)
})
```


```{r}
mx <- st_read(dsn = "shapefiles", layer="muni_2015gw", quiet = T)
```

```{r}
colnames(mx)
```

## Filter and plot with ggplot

```{r}
mx %>% 
  # filter(NOM_ENT %in% c("Quintana Roo", "Yucatán", "Campeche")) %>% 
  ggplot(aes(fill = NOM_MUN)) +
  geom_sf() +
  theme_bw() +
  theme(legend.position = "None")
```

## Filter, merge, plot

```{r}
filter(mx, NOM_ENT %in% c("Baja California", "Baja California Sur", "Sonora", "Sinaloa")) %>% 
  st_union() %>% 
  st_sf(geometry = .) %>% 
  ggplot() +
  geom_sf(fill = "gray") +
  theme_bw()
```

# `raster` package

## Load west coast coastline

```{r}
west <- sf::read_sf(dsn = "shapefiles", layer = "wc_regions_clean")
```

## Specify all raster images

```{r}
sst_files = list.files('rasters', pattern='average_', full.names = T)
```

## Plot one of them

```{r}
r <- raster(sst_files[1])

plot(r)
```

## Histogram of sst of that image

```{r}
hist(r)
```


```{r}
sstStack <-  stack(sst_files)

plot(sstStack)
```

## Raster calculations

I now pass the stacked raster to `calc`, and tell it to calculate the mean, remove the NAs, and substract 273.15 to convert from K to C.

```{r}
sstAvg <-  calc(sstStack, fun = function(x){mean(x, na.rm=T)-273.15})

plot(sstAvg)
```


## In a tidy way

```{r}
stack(sst_files) %>% 
  calc(., fun = function(x){mean(x, na.rm=T)-273.15}) %>% 
  plot()
```

```{r}
stack(sst_files) %>% 
  calc(., fun = function(x){mean(x, na.rm=T)-273.15}) %>% 
  plot()

stack(sst_files) %>% 
  calc(., fun = function(x){mean(x, na.rm=T)-273.15}) %>% 
  contour(asp = 1, add = T)
```

## Load net primary production

```{r}
npp = raster('rasters/annual_npp.tif')

plot(npp)
```

## See extent and projections for each raster

```{r}
sstAvg

npp
```

## Re-project the npp raster to longlat

```{r}
nppProj <-  projectRaster(npp, crs = ('+proj=longlat'))

plot(nppProj)
```

## Crop so that we focus on the US west coast

```{r}
nppCrop = crop(nppProj, sstAvg) #crop nppProj to the extent of sstAvg
plot(nppCrop)
```

## Resample to get a better resolution

```{r}
npp_res = resample(nppCrop, sstAvg, method='ngb')#,progress='text') # Using nearest neighbor
npp_bil = resample(nppCrop, sstAvg, method='bilinear')#,progress='text') # Using bilinear

plot(npp_res)
plot(npp_bil)
```

## Do this the tidy way

```{r}
projectRaster(npp, crs=('+proj=longlat')) %>%
  crop(sstAvg) %>%
  resample(sstAvg, method='ngb') %>%  #,progress='text')
  plot()
```


## Check the new data

```{r}
npp_res
```


## Generate a new SST raster with suitable cells

```{r}
sstPref <- sstAvg
sstPref[sstPref < 12 | sstPref> 18] <- NA

plot(sstPref)

```

## Using the `NA`s, I can assign suitable areas

All cells that are not `NA`s are suitable for the fish:

```{r}
sstPref[!is.na(sstPref)] <- 1

plot(sstPref, col = "coral")
```

## Now for PP

Using reclasify

```{r}
# values between X and Y, get Z
nppBin <- reclassify(npp_res, c(-Inf, 2.6, NA,
                                2.6, 3, 1,
                                3, Inf, NA))

plot(nppBin)
```

## Combine both rasters with `overlay()`

```{r}
overlay(sstPref, nppBin, fun = function(x, y){x*y}) %>% 
  plot(col = "coral")

plot(west, add = T)
```

## Use the shapefile as a mask

```{r}
west_sp <- as(west, "Spatial")

overlay(sstPref, nppBin, fun = function(x, y){x*y}) %>% 
  mask(west_sp) %>% 
  plot()

plot(west, add = T)

```


```{r}
mask(sstAvg, west_sp) %>% 
  plot()

plot(west, add = T)
```


