---
title: "R for scientific computing"
author: "Juan Carlos Villaseñor-Derbez"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_notebook:
    code_folding: hide
    fig_caption: yes
    toc: yes
    toc_collapse: no
    toc_float: yes
---


# Load libraries

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
})
```


# Play with the titanic dataset

## Read the data

```{r}
train <- read.csv(file = "titanic/data/train.csv", stringsAsFactors = F)
test <- read.csv(file = "titanic/data/test.csv", stringsAsFactors = F)
```

## Prepare a training and a testing dataset

```{r}
survival <- select(train, Survived, PassengerId)

train <- mutate(train, Survived = NULL) %>% 
  mutate(is.train = 1)

test <- mutate(test, is.train = 0)

allData <- rbind(train, test)
```

```{r}
totNA <- function(x){sum(is.na(x))}

naCols <- allData %>%
  summarize_all(totNA)

naCols
```

## Obtain titles of the passengers

```{r}
titles <- character(length = nrow(allData))

for (i in 1:nrow(allData)){
  temp <- str_split(allData$Name[i], "\\.")[[1]][1]
  titles[i] <- str_trim(str_split(temp, ",")[[1]][2], "left")
}

allData %<>%
  mutate(title = as.factor(titles)) %>% 
  select(-Name)
```

```{r}
group_by(allData, title) %>% 
  count()
```

## Fill missing information by Title

### Calculate median age by title

```{r}
(medAge <- allData %>% 
  group_by(title) %>% 
  summarize(medAge = median(Age, na.rm = T)))
```

mklink "index.html" "index.nb.html"

### Join median anges to the main dataset, and replace `NA` values with median age

```{r}
allData %<>%
  left_join(medAge, by = "title") %>% 
  mutate(Age = ifelse(is.na(Age), medAge, Age))
```

## Look at the most commonly embarked place

```{r}
(mostCommonEmbarked <- allData %>% 
  group_by(Embarked) %>% 
  summarize(nEmb = n()) %>% 
   arrange(desc(nEmb)) %>% 
   slice(1))
```

And now, instead of running everything else I just source a pre-defined script:


# Run R code on Knot cluster


!/bin/bash

PBS-1 nodes=1:ppn=12

PBS-1 walltime=01:00:00

PBS -N MonteCarlos

PBS -V

cd $PBV_O_WORKDIR

Rscript --vanilla montecarlo.R > output
